---
title: "multiLME-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{multiLME-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning = FALSE, echo = TRUE, results = 'hide'}
devtools::install_github("ERandrinopoulou/multiLME", dependencies = TRUE, force = TRUE)

library(JMbayes2)
library(jagsUI)
library(multiLME)
library(lattice)
```

Let's assume that we want to fit the following multivariate mixed-effects model:

$$
y_{i2}(t) = x_{i2}^\top (t) \beta_2 + z_{i2}^\top (t) b_{i2} + \epsilon_{i2} = m_{i2}(t) + \epsilon_{i2}(t),\\
y_{i1}(t) = x_{i1}^\top (t) \beta_1 + z_{i1}^\top (t) b_{i1} + \epsilon_{i1} = m_{i1}(t) + \alpha m_{i2}(t) + \epsilon_{i1}(t)
$$

In particular, in includes the underlying value of outcome 2 in the mixed-effects model of outcome 1.

First we simulate some data:
```{r simulate, cache = TRUE}
n = 100 

K <- 10  # number of planned repeated measurements per subject, per outcome
t.max <- 10 # maximum follow-up time

################################################
# parameters for the linear mixed effects model 1
betaa1 <- c("(Intercept)" = 3.01, "Time1" = 0.47, "Group" = -2)
sigma1.y <- 1.017671  # measurement error standard deviation

# parameters for the linear mixed effects model 
betaa2 <- c("(Intercept)" = 2.15, "Time1" = 0.91)
sigma2.y <- 1  # measurement error standard deviation

# association parameter
alphaa <- 0.9 # association parameter - value

# Variance-covariance matrix for random effects
invDvec <- c(1, 0, 0, 0,
             0, 1, 0, 0,
             0, 0, 1, 0,
             0, 0, 0, 1)

invDmat <- matrix(invDvec, 4, 4)
D <- solve(invDmat)

################################################
# design matrices for the longitudinal measurement model
times <- c(replicate(n, c(0, sort(runif(K-1, 0, t.max))))) 
group <- sample(c(0,1), n, replace = TRUE)

DF <- data.frame(year = times, drug = factor(rep(group, each = K)))

X1 <- model.matrix(~ year + drug, data = DF)
Z1 <- model.matrix(~ year, data = DF)

X2 <- model.matrix(~ year, data = DF)
Z2 <- model.matrix(~ year, data = DF)

#simulate random effects
library(MASS)

b <- mvrnorm(n, rep(0, nrow(D)), D)

# simulate longitudinal responses
id <- rep(1:n, each = K)

eta.y2 <- as.vector(X2 %*% betaa2 + rowSums(Z2 * b[id, 3:4])) 
y2 <- rnorm(n * K, eta.y2, sigma2.y)

eta.y1 <- as.vector(X1 %*% betaa1 + rowSums(Z1 * b[id, 1:2])) + alphaa * eta.y2 
y1 <- rnorm(n * K, eta.y1, sigma1.y)

dat <- DF[, ]
dat$id <- id
dat$y1 <- y1
dat$y2 <- y2

names(dat) <- c("year", "drug", "id", "y1", "y2")

dat.id <- dat[tapply(row.names(dat), dat$id, tail, 1), ]

data <- dat
data.id <- dat.id
```

Then, we can run the multivariate model as follows:
```{r model, cache = TRUE}
formulas <- list(y1 ~ year + drug + (year | id),
                 y2 ~ year + (year | id))
families <- c("gaussian", "gaussian")

  
fit <- mv_lme(formulas = formulas, data = data, 
                     families = families, 
                     hc = FALSE,
                     corr_RE = FALSE,
                     parameters.b = FALSE,
                     assoc = TRUE,
                     assoc_from = 2,
                     assoc_to = 1,
                     extraForm = NULL)
```

A detailed explanation of the arguments is presented below.
formulas: indicate the formula of the longitudinal models we want to run (without the association)
data: data
families: the families of the outcomes (works for "guassian")
hc: TRUE/FALSE whether hierachical centering should be applied (TRUE option under construction)
predicted: TRUE/FALSE whether the predicted values should be saved
corr_RE: whether correlation between the random effects should be assumed via ac common variance-covariance matrix
engine: only JAGS is supported for now
parameters.b: whether the random effects parameters should be saved
assoc: whether an association parameter between the longitudinal outcomesshould be assumed
assoc_from: indicate which longitudinal outcome should be used as covariate
assoc_to: indicate the main longitudinal model that includes the other longitudinal outcomes as covariates
extraForm: indicate the form of association (e.g., if NULL then the value is assumed, see below more examples)
extraForm_tr: whether a trancated structure should be assumed
norm_area: TRUE/FALSE whether the normalized area should be used (Note: for prediction this is under construction)
time_var: time variable

To obtain results run for example fit$postMeans. 


Let's now assume that we want to obtain individualized predictions:

```{r predictions, cache = TRUE}
object = fit
newdata_all = data[data$id == 1, ]
newdata <- newdata_all[1:6,]

res_pred <- DynPred_mv_lme(object = fit, newdata = newdata, families, hc,
                           level = 0.95, IdVar = "id", timeVar = "year",
                           M = 200, times = NULL,
                           assoc = TRUE,
                           assoc_from = assoc_from,
                           assoc_to = assoc_to,
                           extraForm = extraForm)



xyplot(res_pred$Mean1 + res_pred$Lower1 + res_pred$Upper1 ~ res_pred$Time,
       xlim = c(0, 10), 
       ylim = c(min(y1), max(y1)), 
       strip = FALSE, xlab = "Year", ylab = "Y1",
       lwd = c(3, 2, 2), lty = c(1, 3, 3), type = "l", col = 1,
       panel = function(x, y,  ...) {
         panel.xyplot(x, y, ...)
         panel.abline(v = max(newdata$year), col = 2, lty = 2)
         panel.points(newdata_all$year, newdata_all$y1, pch = 16, cex = 0.8, col = "blue")
         })
```
